\title{The size of things I}
\subtitle{Structure building}
\author{Sabine Laszakovits and Zheng Shen}
\renewcommand{\lsSeries}{ogs}%use series acronym in lower case
\renewcommand{\lsSeriesNumber}{}
\renewcommand{\lsID}{323}
\renewcommand{\lsISBNdigital}{978-3-96110-320-1}
\renewcommand{\lsISBNhardcover}{978-3-98554-016-7}
\BookDOI{10.5281/zenodo.5140049}

\BackBody{This book focuses on the role size plays in grammar. Under the umbrella term \textit{size} fall the size of syntactic projections, the size of feature content, and the size of reference sets. The contributions in this first volume discuss size and structure building.

The most productive research program in syntax where size plays a central role revolves around clausal complements. Part 1 of Volume I contributes to this program with papers that argue for particular structures of clausal complements, as well as papers that employ sizes of clausal complements to account for other phenomena. The papers in Part 2 of this volume explore the interaction between size and structure building beyond clausal complements, including phenomena in CP, vP, and NP domains. The contributions cover a variety of languages, many of which are understudied.

The book is complemented by Volume II which discusses size effects in movement, agreement, and interpretation.}

% title font on cover
\renewcommand{\lsCoverTitleFont}[1]{%
    \sffamily\addfontfeatures{Scale=MatchUppercase}%
    \fontsize{46pt}{15mm}\selectfont #1}

\typesetter{Sabine Laszakovits, Joey Lim, Meghan Lim, Justin Ong, Zheng Shen}
\proofreader{Sabine Laszakovits, Joey Lim, Meghan Lim, Justin Ong, Zheng Shen}
